A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-01-18 paper/draft.tex
continue drafting municipal employment and patronage

Municipal employment in Brazil is under local jurisdiction, with personnel appointments under the exclusive authority of the executive branch. Salaries, contract modalities and terminations are also under municipal jurisdiction. There are multiple forms of contract available, a permanent contract (\emph{estatut\'{a}rio}), a regular contract (\emph{CLT}) and temporary hires. Personnel expenditures cannot exceed a ceiling of 60 percent of the local budget, but as long as the ceiling is respected, other levels of government cannot interfere with local personnel decisions.

Municipal discretion leaves ample room for patronage, whether it be to support mayoral coalitions as outlined in the first paper of this dissertation or to reward contributors and party loyalists \citep{colonnelli2018patronage,brollo2017victor}. These politically motivated appointments are particularly prominent in high-level positions, the so called \emph{cargos de confian\,{c}a}, local ministerial positions that are both high in compensation and provide access to decisions over key public services such as transportation and health.

>>>>

Municipal employment in Brazil is under local jurisdiction, with personnel appointments under the exclusive authority of the executive branch. Salaries, contract modalities and terminations are also under municipal jurisdiction. There are multiple forms of contract available, a permanent contract (\emph{estatut\'{a}rio}), a regular contract (\emph{CLT}) and temporary hires. Personnel expenditures cannot exceed a ceiling of 60 percent of the local budget, but as long as the ceiling is not exceeded, other levels of government cannot interfere with local personnel decisions.

There are no civil service laws regulating municipal employment. With the exception of permanent contracts, municipal employees are subject to the same labor laws as private sector workers. The lack of a civil service system at the local level has been the subject of extensive research \citep{souza2004governos}, and the high turnover that are associated with municipal employment have been widely documented \citep{akhtari2017political}. Labor unions do exist, in particular in the educational sector, but these are regional in focus and concentrated in metropolitan areas. According to the latest education census (\emph{Censo do Magist\'{e}rio}), approximately 11 percent of educational staff was unionized.

Municipal discretion leaves ample room for patronage, whether it be to support mayoral coalitions as outlined in the first paper of this dissertation or to reward contributors and party loyalists \citep{colonnelli2018patronage,brollo2017victor}. These politically motivated appointments are particularly prominent in high-level positions, the so called \emph{cargos de confian\,{c}a}, local ministerial positions that are both high in compensation and provide access to decisions over key public services such as transportation and health. Regarding the nexus between party membership and public sector employment, \citet{brollo2017victor} finds that party members who are politically aligned with the winning mayor are 30 percent more likely to receive public sector employment than their runner-up counterparts.

These empirical findings suggest a clear political nexus between employment into local bureaucracies and political motives in Brazilian municipalities. This paper aims to provide an extensive treatment of the differences in the observable characteristics of party members - as opposed to their non-party members.
